# Exploratory analysis of the data

```{r}
library(tidyverse)
```

Read in primary data file

```{r}
hansard_2006_df <- 
    'data/sure start coverage (hansard 2006) (checked).csv' %>%
    read_csv()

```
## Univariate analysis

```{r}
hansard_2006_df %>% summary 
hansard_2006_df$hansard_2006_coverage %>% 
    quantile(seq(0, 1, 0.1))
```



## Checking against 2009 historical data 

We have seperate repo that collect historical Sure Start information from 2009. We can directly down the data and see whether all wards are covered.

```{r download}
ss_2009_url <- 
    'https://raw.githubusercontent.com/sure-start-ni-evaluation/sure-start-historical-website/refs/heads/main/data/national%20archive%20to%20nimdm%20wards%20(checked).csv'

ss_2009_df <- 
    ss_2009_url %>% read_csv()

```

We can do a full join of the data and see which areas are in both lists 

```{r set-analysis}

ss_2009_df <- ss_2009_df %>% mutate(ss2009 = T)
hansard_2006_df <- hansard_2006_df %>% mutate(ss2006 = T)

full_join_df <-
    hansard_2006_df %>%
    full_join(ss_2009_df, by = 'bestCode')

full_join_df %>% summary

full_join_df %>% 
  filter(is.na(ss2009))
```

Only 1 area in the 2006 list were not in the 2009 list (Camlough). Camlough (in Amagh) appears instead of Carnlough (in Antrim) by 2011. This is due to a data entry error in the 2009 data that was correct later. The responsible Sure Start is South Amagh Sure Start.







## Relationship with multiple deprivation index 

The earliest index of multiple deprivation in NI is from 2001 ([NIMDM 2001](https://www.nisra.gov.uk/publications/northern-ireland-multiple-deprivation-measure-2001-nimdm2001)). The NIMD 2005 was used to decided the expansion of Sure Start.

```{r read in NIMDM}

nimdm2005_df <-
    'data/raw/NIMDM_2010_Results_Ward_0.xls' %>%
      readxl::read_xls(sheet = 2)

nimdm2001_df <-
    'data/raw/nimdm2001.xlsx' %>%
      readxl::read_xlsx(sheet = 2)

```

### Check and join 2001 (matches only)

```{r joindata_2001}

nimdm2001_df <-
    nimdm2001_df %>%
    transmute(
        mdm_score_2001 = `Multiple Deprivation Measure Score`,
        ward_nm = `Ward Name` %>% tolower,
        lgd_nm = `LGD Name` %>% tolower
    )    

## remove ducpliated ward nms (prevent matchign on just first)

nimdm2001_df <- 
    nimdm2001_df[!(nimdm2001_df$ward_nm %>% duplicated),]



joined_hansard_df_2001 <-
    nimdm2001_df %>% 
    left_join(
        hansard_2006_df %>%
        mutate(bestMatch_lower = tolower(bestMatch)), 
        by = c(ward_nm = 'bestMatch_lower')
    ) 


## replace NAs
joined_hansard_df_2001 %>% summary # 97 were found
joined_hansard_df_2001 <-
 joined_hansard_df_2001 %>% 
 replace_na(
    list(hansard_2006_coverage = 0)
 )

```
Let's look at plots of coverage by nimdm 

```{r nimdm2001 plot}

joined_hansard_df_2001 %>%
    ggplot(aes(x = mdm_score_2001, y = hansard_2006_coverage)) +
    geom_point(position = position_jitter(height = 3)) # jitter to improve read

```

Conclusion: There is no real rhyme or reason linking mdm score/rank to SureStart 2006 areas.
